---
title: "7W44_web-scraping"
author: "Ellen Harpøth Lauersen"
date: "2022-11-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
library(rvest)
library(dplyr)
library(tidyr)
library(stringr)
library(janitor)
```


You can choose between three tasks and I have chosen task 2:

_2) use the rvest library to scrape data of your interest (football statistics in Wikipedia?, gender representatives in different governments? global population by country in https://www.worldometers.info/world-population/population-by-country/ )_

I have chosen to use statistics on U.S. Navy aircraft found on wikipedia: https://en.wikipedia.org/wiki/List_of_currently_active_United_States_naval_aircraft

I started out with an attempt at a function, but it wasn't working. The problem turned out be that the website had not one but five tables. I tried specifying which exact table I wanted to harvest, but I was unsuccessful (I carried all this out in a different non-markdown file).


```{r scrape-function, eval=FALSE, echo=TRUE}

fleet_scrape <- function(website){
  
  # scraping website
  url <- read_html(website)
  
  # creating table named as specified with function-input
  fleet_table <- url %>%
    html_nodes("wikitable") %>%
    html_table()
  fleet_table <- do.call(cbind,unlist(fleet_table, recursive = FALSE))
}

```


Trying to call it up like this, I realised it hadn't worked.


```{r, eval=FALSE, echo=TRUE}

USnavy_airfleet <- fleet_scrape("https://en.wikipedia.org/wiki/List_of_currently_active_United_States_naval_aircraft")

USnavy_airfleet %>%
  head()

```


Instead I decided to drop the function and do every step of the scraping process seperately, so I could find out exactly where the probem was arising. It turned out that if I harvested all five tables at once but then just "extracted" the particular table I wanted into its own object, then I could get it to work.
As seen here:

```{r fleet-code}

# making a url-object
url <- "https://en.wikipedia.org/wiki/List_of_currently_active_United_States_naval_aircraft"

# scraping the website
url_html <- read_html(url)

# scraping the tables
navy_table <- url_html %>%
  html_nodes("table") %>%
  html_table()

# inspecting the tables
head(navy_table)

# identifying the type of my "table" (which is a tibble)
str(navy_table)

# extracting the first table (which is the table of aircraft (and helicopters))
navy_aircraft <- navy_table[[1]]

```

Just inspecting the first part of my table, I can ssee that this worked.

```{r fleet_table, echo=TRUE, include=TRUE}

head(navy_aircraft)

```


This could perhaps be made into a function, but I don't think it would have the desired reusability as the table you need to extract might be different each time you use it on a different website. For this reason I won't bother making a function out of it.

Upon further inspection though the table has some weird rows that identify "sub-tables" within the table (combat aircraft, electronic warfare, maritime patrol, transport, helicopter). I would attempt to fix this except it's not a required part of the task, I don't really know how to do it, and I don't have the time to figure it out. But this would be somethin necessary to do for a neater dataframe.